---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Reading the csv file
```{r 1_1}
dat0 <- read.csv("activity.csv")
```
Transforming Date variable from character to date format
```{r 1_2}
dat0$date <- as.Date(dat0$date,"%Y-%m-%d")
```

## What is mean total number of steps taken per day?
Aggregating total no of steps over day to new data frame
```{r 2_1}
dat1 <- aggregate(steps~date,dat0,sum,na.rm=TRUE)
```
Plotting histogram of total number of steps per day
```{r 2_2}
hist(dat1$steps,main="Histogram of Total Steps per Day",xlab="No of Steps")
```

Mean and median of total number of steps per day
```{r 2_3}
mean(dat1$steps)
median(dat1$steps)
```

## What is the average daily activity pattern?
Aggregating mean no of steps over interval to new data frame
```{r 3_1}
dat2 <- aggregate(steps~interval,dat0,mean,na.rm=TRUE)
```
Plotting time series plot of average daily activity pattern
```{r 3_2}
plot(dat2$interval,dat2$steps,main="Average Daily Activity Pattern",
     xlab="5-Minute Interval",ylab="Average No of Steps",type="l")
```

Getting interval which contains max no of steps
```{r 3_3}
dat2$interval[which(dat2$steps==max(dat2$steps))]
```

## Imputing missing values
Getting total number of missing values
```{r 4_1}
sum(is.na(dat0))
```
Using mean no of steps per 5-minute interval for missing values imputation
```{r 4_2}
dat3 <- dat0
dat3$steps[is.na(dat3$steps)] <- dat2[match(dat0$interval[is.na(dat0$steps)],
                                            dat2$interval),"steps"]
```
Aggregate total no of steps over day to new data frame
```{r 4_3}
dat4 <- aggregate(steps~date,dat3,sum,na.rm=TRUE)
```
Plotting histogram of total steps per day after imputation
```{r 4_4}
hist(dat4$steps,main="Histogram of Total Steps per Day",xlab="No of Steps")
```

Getting mean and median of total steps per day, the values are close to mean and median 
before imputation as my imputation strategy is using no of steps for 5-minute
interval averaged over day
```{r 4_5}
mean(dat4$steps)
median(dat4$steps)
```

## Are there differences in activity patterns between weekdays and weekends?
Creating new factor variable which indicates weekday or weekend
```{r 5_1}
dat3$daytype <- ifelse(weekdays(dat3$date) %in% c("Saturday","Sunday"),
                       "weekend","weekday")
dat3$daytype <- as.factor(dat3$daytype)
```
Aggregating mean no of steps over interval and daytype to a new data frame
```{r 5_2}
dat5 <- aggregate(steps~interval+daytype,dat3,mean,na.rm=TRUE)
```
Plotting time series plot of average daily activity pattern by weekday and weekend
```{r 5_3}
library(lattice)
xyplot(steps ~ interval | daytype, data=dat5, type="l", layout=c(1,2),
       xlab="Interval", ylab="No of Steps")
```
